Timid tides or tall tails? \\ Investigating the outer reaches of the Sculptor and Ursa Minor dwarf galaxies

### 2025-04-08 thesis/introduction.tex
thesis updates to methods

@@

\begin{figure}
\centering
\includegraphics[width=5.41667in,height=5.41667in]{figures/scl_des_dr2.png}
\caption[scl\_image]{Sculptor image from DES DR 2 via
HiPS2FITS.}\label{fig:scl_image}
\end{figure}

@@

While Scl has a relatively large pericentre (greater than 50kpc),
\citet{Sestito+23} detect that the galaxy likely has an excess of stars
in the outskirts (past about 60 arcminutes). This excess is perhaps one
of the clearest indications that Scl may be affected by tides. Here, our
goal is to determine if under a \(\Lambda\) CMD paradigm with DM-only
simulations, tidal effects of the Milky Way (or LMC) may indeed be
consistent with these observations, or if these observations may reveal
instead a extended stellar ``halo'' or second component of the
galaxy---illustrating a complex formation for the galaxy.

Figure: Image of Sculptor from DES.

@@

\begin{table*}[t]
\centering
\caption{Measured properties of Sculptor}
\label{tbl:Measured-properties-of-Sculptor}
\begin{tabular}{lll}
\toprule
parameter & value & Source\\
\midrule
$\alpha$ & $15.0183 \pm 0.0012$˚ & M+18\\
$\delta$ & $-33.7186 \pm 0.00072$˚ & M+18\\
distance & $83.2 \pm 2$ kpc & Tran+22\\
$\mu_\alpha \cos \delta$ & $0.099 \pm 0.002 \pm 0.017$ mas yr$^{-1}$ & MV20a\\
$\mu_\delta$ & $-0.160 \pm 0.002_{\rm stat} \pm 0.017_{\rm sys}$ mas yr$^{-1}$ & MV20a\\
RV & $111.4 \pm 0.2$ & This work\\
$\sigma_v$ & $9.61\pm0.16$ & This work\\
$R_h$ & $12.33 \pm 0.05$ arcmin & MV20*\\
$R_h$ & $8.69 \pm 0.2$ arcmin & this work?\\
ell & $0.36 \pm 0.01$ & M+18\\
PA & $92\pm1$ & M+18\\
$M_V$ & $-10.82\pm0.14$ & M+18\\
\bottomrule
\end{tabular}
\end{table*}

@@

\begin{figure}
\centering
\includegraphics[width=5.41667in,height=5.41667in]{figures/umi_image.jpg}
\caption[Ursa Minor Image]{Ursa Minor Dwarf. Composite Public
data/Giuseppe Donatiello. Updated image in 2022.
https://www.flickr.com/photos/133259498@N05/page8}\label{fig:umi_image}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  Cosmology foundations
\item
  NFW plot (density or energy), maybe borrow from paper?

  \begin{itemize}
  \tightlist
  \item
    Explain origin of NFW
  \end{itemize}
\item
  Dwarf galaxy formation, halos
\item
  Evolution under tidal field
\item
\end{itemize}

@@

>>>>

\subsection{Cosmological context}\label{cosmological-context}

\begin{figure}
\centering
\includegraphics{/Users/daniel/Library/Application Support/typora-user-images/image-20250408091622850.png}
\caption[Cosmological Power Spectrum]{figure 1 from
\citet{bechtol+2022}.}
\end{figure}

Figure: Density profiles of comological simulated halos, matching
approximantly the NFW formula.

@@

\begin{figure}
\centering
\includegraphics[width=5.41667in,height=5.41667in]{figures/scl_des_dr2.png}
\caption[Picture of Sculptor]{Sculptor image from DES DR 2 via
HiPS2FITS.}\label{fig:scl_image}
\end{figure}

@@

While Scl has a relatively large pericentre (greater than 50kpc),
\citet{sestito+2023a} detect that the galaxy likely has an excess of
stars in the outskirts (past about 60 arcminutes). This excess is
perhaps one of the clearest indications that Scl may be affected by
tides. Here, our goal is to determine if under a \(\Lambda\) CMD
paradigm with DM-only simulations, tidal effects of the Milky Way (or
LMC) may indeed be consistent with these observations, or if these
observations may reveal instead a extended stellar ``halo'' or second
component of the galaxy---illustrating a complex formation for the
galaxy.

@@

\begin{table*}[t]
\centering
\caption{Measured properties of Sculptor}
\label{tbl:Measured-properties-of-Sculptor}
\begin{tabular}{lll}
\toprule
parameter & value & Source\\
\midrule
$\alpha$ & $15.0183 \pm 0.0012$˚ & M+18\\
$\delta$ & $-33.7186 \pm 0.00072$˚ & M+18\\
distance & $83.2 \pm 2$ kpc & Tran+22\\
$\mu_\alpha \cos \delta$ & $0.099 \pm 0.002 \pm 0.017$ mas yr$^{-1}$ & MV20a\\
$\mu_\delta$ & $-0.160 \pm 0.002_{\rm stat} \pm 0.017_{\rm sys}$ mas yr$^{-1}$ & MV20a\\
RV & $111.2 \pm 0.2$ & arroyo-polonio+24\\
$\sigma_v$ & $9.7\pm0.2$ & arroyo-polonio+24\\
$R_h$ & $12.33 \pm 0.05$ arcmin & MV20*\\
$R_{h,inner}$ & $8.69 \pm 0.2$ arcmin & this work?\\
ell & $0.36 \pm 0.01$ & M+18\\
PA & $92\pm1$ & M+18\\
$M_V$ & $-10.82\pm0.14$ & M+18\\
\bottomrule
\end{tabular}
\end{table*}

@@

\begin{figure}
\centering
\includegraphics[width=5.41667in,height=5.41667in]{figures/umi_image.jpg}
\caption[Picture of Ursa Minor]{Ursa Minor Dwarf. Composite Public
data/Giuseppe Donatiello. Updated image in 2022.
https://www.flickr.com/photos/133259498@N05/page8}\label{fig:umi_image}
\end{figure}

@@

\begin{itemize}
\tightlist
\item
  Cosmology foundations

  \begin{itemize}
  \tightlist
  \item
    power spectrum plot
  \end{itemize}
\item
  NFW plot (density or energy), maybe borrow from paper?

  \begin{itemize}
  \tightlist
  \item
    Explain origin of NFW
  \end{itemize}
\item
  Dwarf galaxy formation, halos
\end{itemize}

N-body DM simulations

\begin{itemize}
\tightlist
\item
  Collisionless Boltzmann equation and meaning of such simulations
\item
  Assumptions \& context \& past work
\item
  Evolution under tidal field
\end{itemize}

@@

\subsection{Introduction to Dark Matter
simulations}\label{introduction-to-dark-matter-simulations}

In this section, we will cover

\begin{itemize}
\tightlist
\item
  How are dark matter simulations conducted
\item
  Interpretations and uncertainties
\item
  Methods
\end{itemize}

\begin{figure}
\centering
\includegraphics{/Users/daniel/thesis/figures/idealized_break_radius.png}
\caption[Break radius validation]{The break radius of the simulations is
set by}\label{fig:idealized_break_radius}
\end{figure}

From this argument, we note that the following properties must be
approximately true for tides to occur:

\begin{itemize}
\tightlist
\item
  Close enough pericentre. The other break radius \(r_J\) implies that
  if the host density is 3x the satellite, stars will be lost
\item
  Corresponding time since last pericentre: If the time since last
  pericentre is not \(\sim\) consistent with an observed break in the
  density profile, then tides
\item
  Halo evolution. As found in \citet{EN2021}, galaxies evolve along well
  defined tidal tracks (assuming spherical, isotropic, NFW halo, which
  may not be true, see \ldots). These tracks tend to ``puff up'' the
  stellar component while also removing dark matter mass, leaving a
  smaller, compacter DM halo with a more extended stellar component.

  \begin{itemize}
  \tightlist
  \item
    This information is mostly related to the statistical initial
    distribution of satellites from cosmology {[}ludlow+2016;
    \citet{fattahi+2018}{]}
  \end{itemize}
\end{itemize}

\subsection{Potentials}\label{potentials}

From the \citet{nfw1996} paper, eqns. 3 \& 4 \[
\frac{\rho}{\rho_c} = \frac{\delta_c}{(r/r_s)(1+r/r_s)^2}
\] where \[
\delta_c = \frac{200}{3}\frac{c^3}{[\ln(1+c)-c/(1+c)]}
\] \(c\) is concentration parameter, \(\rho_c\) is the critical density
of the universe, and \(r_s\) is the characteristic scale length of the
halo.

The NFW halo is sometimes described by \(M_{200}\). \(r_{200}\) is the
radius at which the mean density of the halo interior to \(r_{200}\) is
200 times the critical density of the universe, and \(M_{200}\) is the
mass contained inside \(r_{200}\). As equations: \[
\rho_{200} = 200\rho_{c} = \frac{M_{200}}{(4\pi/3) r_{200}^3}
\]

\[
r_{200} = \sqrt[3]{\frac{1}{200}\frac{3M_{200}}{4\pi \rho_{\rm c
}}}
\]

\[
M_{200} = \frac{4\pi}{3} r_{200}^3\ \rho_{200}
\]

\(M_{200}\) is also sometimes called the virial mass of the halo.
\(r_{200}\) is directly related to \(r_s\) by \[
r_{200} = c\,r_s
\]

Another useful definition is \[
A(x) \equiv \log (1+x) - \frac{x}{1+x}.
\]

We will also define a dimensionless radius \[
x \equiv r/r_s.
\]

A simple substitution to the definition gives

\[
\rho(x) =  \frac{\rho_s/3}{x\ (1+x)^2}
\] where \[
\rho_s = \frac{c^3}{A(c)} \rho_{200}
\] and \(A(c)\) is as above. The characteristic density can also be
written in terms of scale mass, \(M_s = M_{200}/{A(c)}\) (see below),
giving \[
\rho_s = \frac{c^3}{A(c)} \frac{M_{200}}{(4\pi/3)\ r_{200}^3}  = \frac{3M_s}{4\pi\, {r_s}^3}
\] Note that the NFW density profile is the same as an alpha-beta-gamma
profile where \(\alpha=\gamma=1\) and \(\beta =3\).

\subsubsection{Circular velocity}\label{circular-velocity}

The circular velocity in terms of
\(v_{200} = \sqrt{G M_{200} / R_{200}}\) is \[
\left(v_{\rm circ}/v_{200}\right)^2 = \frac{A(x)/x}{A(c)/c},
\]

or in terms of \(M_s\) and \(r_s\), \[
v_{\rm circ}^2 = \frac{G M(r)}{r} = \frac{G M_s A(r/r_s)}{r}.
\] Another parameterization of the NFW profile is in terms of the
maximum circular velocity \(v_{\rm circ}^{\rm max}\) and the radius at
which it is reached \(r_{\rm circ}^{\rm max}\). Given the scale radius,
\[
r_{\rm circ}^{\rm max} = \alpha\ r_s
\]

where \(\alpha\approx2.16258\), and \(v_{\rm circ}^{\rm max}\) can be
found from either of the equations above.

### 2025-04-08 thesis/methods.tex
thesis updates to methods

We use Agama \citep{vasiliev} for potential specification and initial
conditions, and Gadget-4 \citep{springle2021} for orbit integration and
dark matter simulations (with a custom patch to integrate with Agama).

@@

Our fiducial potential is as described in \citet{EP2020}. This potential
is an analytic approximation of \citet{McMillan2011}, consisting of a
stellar Bulge, disk, and a dark matter NFW halo. The specific parameters
are

\subsubsection{Bulge}\label{bulge}

Hernquist potential with \(r_s=1.3\), mass = 2.1.

@@

Density Profile \[
\rho(r) = \frac{M}{2\pi} \frac{a}{r} \frac{1}{(r+a)^3}
\]

\subsubsection{Disk}\label{disk}

@@

for convenience, we define the helper variables,
\(S = \sqrt{z^2 + b^2}\) and \(D = \sqrt{R^2 + (a+S)^2}\), so the
potential simplifies to \[
\Phi(R, z) = \frac{-GM}{D}
\]

\[
\rho = \frac{b^2 M}{4\pi} \frac{aR^2 + (a+3r)(a + r)^2}{S^3 D^5}
\]

\subsubsection{Halo}\label{halo}

Mvir=115, r=20.2, c=9.545

@@

Variations to the potential of the inner disk (exclusion of a bar)
should minimally affect our results as no orbit we consider reaches less
than \textasciitilde15 kpc of the MW centre.

\textbf{Figure:} Circular velocity profile of \citet{EP2020} potential.

@@

To estimate the orbits of a dwarf galaxy, we perform a Monte Carlo
sampling of the present-day observables. We integrate each sampled
observable back in time for 10 Gyr in Gadget as massless point
particles.

We use the astropy v4 galactocentric frame. This frame assumes
\ldots\ldots{}

@@

We use Agama (cite) to generate initial conditions. We initially assume
galaxies are described by an NFW dark matter potential and the stars are
merely collisionless tracers embedded in this potential

@@

\begin{itemize}
\tightlist
\item
  \(\Psi\) is taken as known from the underlying assumed analytic dark
  matter potential.
\item
  \(\rho_\star\) is
\end{itemize}

\subsubsection{Stellar profiles}\label{stellar-profiles}

We consider several different stellar profiles in this work.

@@

The 2D exponential stellar profile is given by \[
\Sigma_\star(R) = A e^{-R / R_s}
\] where \(A\) is some normalization and \(R_s\) is the exponential
scale radius. The 3D deprojected profile is found through abel inversion
(e.g.~Rapha\ldots), i.e. \[
\rho_\star (r) = \int dR = xxx
\]

@@

@@

See \citet{power+2003} for a discussion of this.

The gravitational softening is based on: \[
h_{grav} = 4 \frac{R_{200}}{\sqrt{N_{200}}}
\]

I in fact find that dividing the softening by a factor of sqrt(10) gives
slightly more precise results with a negligible increase in compute
time.

@@

Here, we describe some convergence tests to ensure our methods and
results are minimally impacted by numerical limitations and assumptions.
See \citet{power2003} for a detailed discussion of various assumptions
and parameters used in N-body simulations.

\subsubsection{Particle Number.}\label{particle-number.}

\subsubsection{Timestepping}\label{timestepping}

\subsubsection{Gravitational methods}\label{gravitational-methods}

\subsubsection{Gravitational accuracy}\label{gravitational-accuracy}

@@

\begin{verbatim}
# -----------------------------------IO---------------------------------------

@@

#---Mem & CPU limits
TimeLimitCPU                86400   # TUNE
CpuTimeBetRestartFile       7200    # TUNE 
MaxMemSize                  2400    # TUNE to system (MB)

@@

#---Output frequency
OutputListOn                0
TimeBetSnapshot             10      # change as needed
TimeOfFirstSnapshot         0       # MATCH to TimeBegin
TimeBetStatistics           10      # change as needed 
NumFilesPerSnapshot         1
MaxFilesWithConcurrentIO    1

# ----------------------------------Gravity-----------------------------------

# MATCH all the below if continuing run

#---Timestep accuracy
ErrTolIntAccuracy           0.01    # MATCH
CourantFac                  0.1     # ignored; for SPH
MaxSizeTimestep             0.5     # MATCH
MinSizeTimestep             0.0

#---Tree algorithm
TypeOfOpeningCriterion      1       # 0: Barnes-Hut, 1: Relative
ErrTolTheta                 0.5     # mostly used for Barnes-Hut
ErrTolThetaMax              1.0     # (used only for relative)
ErrTolForceAcc              0.005   # (used only for relative)

@@

#---Gravitational Softening
SofteningComovingClass0      # FILL IN
SofteningMaxPhysClass0       0      # ignored; for cosmological
SofteningClassOfPartType0    0
SofteningClassOfPartType1    0

#--------------------------------Misc-------------------------------------

>>>>

We use Agama \citep{agama} for potential specification and initial
conditions, and Gadget-4 \citep{gadget4} for orbit integration and dark
matter simulations (with a custom patch to integrate with Agama).

@@

Our fiducial potential is as described in \citet{EP2020}. This potential
is an analytic approximation of \citet{McMillan2011}, consisting of a
stellar Bulge, disk, and a dark matter NFW halo.
fig.~\ref{fig:v_circ_potential} plots the circular velocity profiles.

@@

\[
\rho(r) = \frac{M}{2\pi} \frac{a}{r} \frac{1}{(r+a)^3}
\] Hernquist potential with \(r_s=1.3\), mass = 2.1.

@@

@@

Variations to the potential of the inner disk (exclusion of a bar)
should minimally affect our results as no orbit we consider reaches less
than \textasciitilde15 kpc of the MW centre. We exclude the mass
evolution of the halo from this analysis. Over 10 Gyr, this would be
fairly significant (\textasciitilde2x in MW mass) but since we want to
determine the upper limit of tidal effects, it is safe to neglect this.

\begin{figure}
\centering
\includegraphics{figures/v_circ_potential.pdf}
\caption{Circular velocity of potential}\label{fig:v_circ_potential}
\end{figure}

\textbf{Figure:} Circular velocity profile of \citet{EP2020} potential.
We also show the MW and LMC potential from \citet{vasiliev2024} for
their L3M11 model.

@@

To estimate the orbits of a dwarf galaxy, we perform a Monte Carlo
sampling of the present-day observables. We integrate each sampled
observable back in time for 10 Gyr in Gadget as massless point particles
outputing every 0.5Myr (with otherwise similar parameters to n-body runs
below.)

To convert from Gaia to galactocentric coordinates, we use the astropy
v4 galactocentric frame. This frame assumes an

\begin{itemize}
\tightlist
\item
  J2000 `ra' = 17h45m37.224s, `dec' = -28°56'10.23'' (appendix),
  `pmracosdec=-3.151±0.018 mas/yr', `pmdec=-5.547±0.026' (Table 2). from
  \citet{reid+brunthaler2004}
\item
  distance = 8.122 ± 0.033 kpc, solar radial velocity = 11 + 1.9 ± 3
  km/s. from \citet{gravitycollaboration+2018}.
\item
  `z\_sun' = 20.8±0.3pc from bennett+bovy2019
\item
  `v\_sun' is slightly more complicated, relying on parameters from
  above and using the procedure described in \citet{drimmel+poggio2018}.
  This results in

  \begin{itemize}
  \tightlist
  \item
    `v\_sun' = {[}-12.9 ± 3.0, 245.6 ± 1.4, 7.78 ± 0.08{]} km/s
  \end{itemize}
\end{itemize}

@@

We use Agama \citep{agama} to generate initial conditions. We initially
assume galaxies are described by an NFW dark matter potential (REF) and
the stars are merely collisionless tracers embedded in this potential.
We adopt a truncation radius of 100 times the scale radius (likely
larger than necessary, but should not affect results.)

@@

\begin{itemize}
\tightlist
\item
  \(\Psi\) is taken as known from the underlying assumed analytic dark
  matter potential.
\item
  \(\rho_\star\) is pre-specified or calculated via abel integration /
  deprojection
\item
  \(f(\epsilon)\) (the energy distribution function) is determined by
  Edington inversion
\end{itemize}

To find the distribution function, we use Eddington inversion (eq.
4-140b in BT87) \[
f({\cal E}) = \frac{1}{\sqrt{8}\, \pi^2}\left( \int_0^{\cal E} \frac{d^2\rho}{d\Psi^2} \frac{1}{\sqrt{{\cal E} - \Psi}}\ d\Psi + \frac{1}{\sqrt{\cal E}} \left(\frac{d\rho}{d\Psi}\right)_{\Psi=0} \right)
\]

where \({\cal E}\) is the binding energy, and \(\Psi\) is the potential
normalized to go to zero at infinity. In practice the right, boundary
term is zero as \(\Psi \to 0\) as \(r\to\infty\), and if
\(\rho \propto r^{-n}\) at large \(r\) and \(\Psi \sim r^{-1}\) then
\(d\rho / d\Psi \sim r^{-n+1}\) which goes to zero provided that
\(n > 1\).

We consider both a Plummer and 2-dimensional exponential (Exp2D) stellar
profile:

@@

The 2D exponential stellar profile is given by \[
\Sigma_\star(R) = A e^{-R / R_s}
\] where \(A\) is some normalization and \(R_s\) is the exponential
scale radius. The 3D deprojected profile is found through abel inversion
(e.g.~Rapha\ldots), i.e. \[
\rho_\star (r) =- \frac{1}{\pi}\int_r^\infty \frac{d\Sigma}{dR} \frac{1}{\sqrt{R^2 - r^2}} dR  = \frac{\Sigma_0}{\pi R_s^2}\,K_0(r/R_s)
\] where \(K_0\) is the 0th order modified Bessel function of the second
type.

@@

A Plummer profile is defined by \[
\Sigma(R) = \frac{M}{4\pi R_s^2} \left(1 + \frac{R^2}{R_s^2}\right)^{-2} ,
\] where \(M\) is the total mass and \(R_s\) is the characteristic scale
radius. The density is \[
\rho(r) = \frac{3M}{4\pi\,R_s^3} \left(1 + \frac{r^2}{R_s^2}\right)^{-5/2}.
\]

@@

We adopt a softening length of \[
h_{\rm grav} = 0.014 \left(\frac{r_s}{2.76\,{\rm kpc}}\right)\left(\frac{N}{10^7}\right)^{-1/2}
\] see appendix for a discussion of this choice, which is similar to the
\citet{power+2003} suggested softening.

\subsubsection{Analysis}\label{analysis}

Shrinking spheres centres inspired by \citet{power+2003}

\begin{itemize}
\tightlist
\item
  Recursively shrink radius by 0.975 quantile and recalculating centroid
  until radius is less than \textasciitilde1kpc
\item
  Remove bound particles (using instantanious potential).
\item
  Use previous snapshot centre and acceleration to predict new centre
  for next snapshot.
\end{itemize}

@@

Here, we describe some convergence tests to ensure our methods and
results are minimally impacted by numerical limitations and assumptions.
See \citet{power+2003} for a detailed discussion of various assumptions
and parameters used in N-body simulations.

\subsubsection{Softening}\label{softening}

\citet{power+2003} suggest the empirical rule that the ideal softening
(balancing integration time and only compromising resolution in
collisional regime) is \[
h_{grav} = 4 \frac{R_{200}}{\sqrt{N_{200}}}
\]

For our isolation halo (\(M_s=2.7\), \(r_s=2.76\)) and with \(10^7\)
particles, this works out to be \(0.044\,{\rm kpc}\).We adpoted the
slightly smaller softening which was reduced by a factor of
\(\sqrt{10}\) based on initial numerical experiments. However, this
choice is likely too small and a larger softening would minimally impact
our results but improve computation speed.

\subsubsection{Particle Number}\label{particle-number}

As discussed in \citet{power+2003}, the region where density is
converged is related to the region which becomes collisionally relaxed
over the time of the universe (so about 5-10 Gyr). (i.e.~where the
collisionless assumption breaks down). The relaxation timescale is given
by \(t_{\rm relax} = t_{\rm circ} N(r) / 8\ln N(r)\) or otherwise, \[
t_{\rm relax}(r) := t_{\rm circ}(r) \frac{N(r)}{8\,\ln N(r)}
= {t_{\rm circ}(r_{200})} \frac{\sqrt{200}}{8} \frac{N(r)}{\ln N(r)} \left(\frac{\bar \rho (r)}{\rho_{\textrm crit}}\right)^{-1/2}
\] This works out to be about 6-10 times (increasing with particle
number) our adopted softening length for NFW halos given our
assumptions. As such, at full resolution, we can only trust density
profiles down to \(10\epsilon\) or about
\(0.14\,{\rm kpc} (r_s/2.76\,{\rm kpc})\), just enough to resolve
stellar density profiles.

Note that by decreasing the truncation radius, the effective particle
number is improved, so future experiments could be less generous with
this parameter to improve computational performance (order 30\% better
resolution).

\begin{figure}
\centering
\includegraphics{figures/iso_converg_num.pdf}
\caption[num convergence]{Numerical convergence test for circular
velocity as a function of log radius for simulations with different
total numbers of particles in isolation. Residuals in lower panel are
relative to NFW. The initial conditions are dotted and the converged
radius is marked by arrows (REF).}
\end{figure}

\subsubsection{Timestepping and force
accuracy}\label{timestepping-and-force-accuracy}

In general, we use adaptive timestepping and relative opening criteria
for gravitational force computations. To verify that these choices and
associated accuracy parameters minimally impact convergence or speed, we
show a few more isolation runs (using only 1e5 particles)

\begin{itemize}
\tightlist
\item
  constant timestep (\ldots), approximantly half of minimum timestep
  with adaptive timestepping
\item
  geometric opening, with \(\theta = 0.5\).
\item
  strict integration accuracy, (facc = \ldots.)
\end{itemize}

\subsubsection{Alternative methods}\label{alternative-methods}

\begin{itemize}
\tightlist
\item
  FMM
\item
  PMM-tree
\item
  Gadget2
\item
  etc.
\end{itemize}

@@

Note that we use code units which assume that \(G=1\) for convenience
and numerical stability. The conversion between code units to physical
units is (for our convention):

\begin{itemize}
\tightlist
\item
  1 length = 1 kpc
\item
  1 mass unit = \(10^{10}\) Msun
\item
  1 velocity unit = 207.4 km/s
\item
  1 time unit = 4.715 Myr
\end{itemize}

Most parameters below are not too relevant or have been discussed or are
merely dealing with cpu and IO details. The changes between simulation
runs primarily affect the integration time, output frequency, and
softening. Otherwise, we leave all other parameters fixed.

\begin{verbatim}
#======IO parameters======

@@

#---Mem & CPU limits
TimeLimitCPU                86400
CpuTimeBetRestartFile       7200
MaxMemSize                  2400

@@

#---Output frequency
OutputListOn                0
TimeBetSnapshot             10
TimeOfFirstSnapshot         0 
TimeBetStatistics           10
NumFilesPerSnapshot         1
MaxFilesWithConcurrentIO    1

#=======Gravity======

#---Timestep accuracy
ErrTolIntAccuracy           0.01
CourantFac                  0.1     # ignored; for SPH
MaxSizeTimestep             0.5
MinSizeTimestep             0.0

#---Tree algorithm
TypeOfOpeningCriterion      1       # Relative
ErrTolTheta                 0.5     # mostly used for Barnes-Hut
ErrTolThetaMax              1.0     # (used only for relative)
ErrTolForceAcc              0.005   # (used only for relative)

@@

#---Gravitational Softening
SofteningComovingClass0      0.044  # HALO dependent
SofteningMaxPhysClass0       0      # ignored; for cosmological
SofteningClassOfPartType0    0
SofteningClassOfPartType1    0

#=======Miscellanius=======
